---
title: "TidyTuesday 2021-06-15 - DuBoisChallenge"
author: "Cristóbal Alcázar"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r toy-tools, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggtext)
library(tidytext)
library(ggwordcloud)
library(patchwork)
library(magick)

theme_set(theme_minimal(base_family = "Roboto Condensed",
                        base_size = 12))
```


```{r download-data}
tweets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-15/tweets.csv')
```

## Visualising #DuBoisChallenge tweets


```{r}
# 0. Data preparation for visualize tweet's activity
# ------------------------------------------------------------------------------
# Highlight the days in which are popular tweets 
popular_tweets <- tweets %>% 
  filter(like_count >= 50) %>% 
  mutate(date = lubridate::as_date(datetime)) %>% 
  select(date, like_count)
 
# Count the tweets per day
tweets_per_day <- tweets %>% 
  mutate(date = lubridate::as_date(datetime)) %>% 
  group_by(date) %>% 
  count(sort = TRUE) %>% 
  ungroup()


# Make a wordcloud
set.seed(42)
top_wordCloud <- tweets %>% 
  select(content) %>% 
  unnest_tokens(word, content) %>% 
  anti_join(stop_words, by = "word") %>% 
  group_by(word) %>% 
  count(sort = TRUE) %>% 
  ungroup() %>% 
  slice(1:50) %>% 
  mutate(angle = 90 * sample(c(0,1),n(), replace=TRUE, prob=c(60,40))) %>% 
  ggplot(aes(label=word, size=n, angle=angle)) +
  geom_text_wordcloud(colour="#c51f5d") +
  scale_size_area(max_size = 8) +
  theme(plot.background = element_rect(color = "#FFFCFC", fill = "#141d26"))
                  


# 1. polar-tweet activty using geom_step [BASE PLOT]
# ------------------------------------------------------------------------------
base_plot <- tweets_per_day %>% 
  filter(date < "2021-05-01") %>% 
  ggplot(aes(date)) +
  # trend following the chaning in the # of tweets
  geom_step(aes(y=n, colour=n),
            size=0.8) +
  expand_limits(y = 120) +
  # time orientation arrow to read the plot (geom_segment + coord_polar = geom_curve):
  geom_segment(aes(x=lubridate::as_date("2021-04-10"), xend=lubridate::as_date("2021-05-01"),
                   y=100, yend=100),
               size=0.05,
               lineend = "round",
               linejoin = "bevel",
               arrow = arrow(length = unit(0.15, "inch")),
               colour = "orange") +
  # inner circle (induce -Inf / Inf)
  geom_segment(aes(x = min(date), y=500, xend=max(date),yend=500),
               colour = "#657786",
               size = 0.7,
               data = filter(tweets_per_day, between(date, "2021-04-15", "2021-05-07"))) +
  # using twitter night-mode colours
  scale_colour_gradient2(low = "white", mid = "#243447", high = "#c51f5d") +
  # introduce tweets with a lot of likes
  geom_jitter(aes(y=1200, size = like_count), 
              shape = 21,
              fill = "#c51f5d",
              colour = "white",
              alpha = .6,
              data = popular_tweets) +
  # use log scale
  scale_y_log10() +
  # make everything polar 
  coord_polar() +
  labs(x = NULL,
       y = NULL,
       title="Visualising #DuBoisChallenge tweets",
       # subtitle="A #TidyTuesday week 8-2021",
       size = "                Tweets with\n           more than X likes:",
       colour = "# of tweets",
       caption = "<br>**Source:** #DuBoisChallenge tweets<br>
                 Graphic: *@vamos_alcazar*") +
  theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        
        plot.margin = margin(25, 50, 15, 25),
        plot.background = element_rect(color = "#FFFCFC", fill = "#141d26"),
        plot.title = element_text(colour = "white", size = 18),
        plot.subtitle = element_text(colour = "white", size = 14),
        plot.caption = element_markdown(colour = "white", lineheight = 1.2),
        
        axis.text = element_text(colour = "white", size=12, vjust=.5),
        axis.text.y = element_blank(),
        
        legend.title = element_text(colour = "white"),
        legend.text = element_text(colour = "white"),
        legend.position = "bottom",
        legend.box = "horizontal"
        ) +
  guides(
         colour = guide_colorbar(title.position = "top",
                                 title.hjust = 0.5),
         size = guide_legend(title.position = "top")
        )

base_plot

# using patchwork to combine base_plot a top_wordCloud into a single visualisation
(base_plot + top_wordCloud)  +
  plot_layout(heights = c(1, .45))

# save plot
ggsave("2021-06-15_duBoisChallenge.png", dpi=320)
```


- using gradient colour
- background greydark
- color blending - Infinity Showcase
- blend colors for overlapping areas
- combining charts
- Create a gap using the data representation with same colour as the background
to create a visual space (Method 4: Hacking the Code / Nadieh Bremer lesson)
- Refactor a donnut chart / histogram / bar chart with polar coord

## EDA

```{r}
# # explore datetime range
tweets %>%
  summarize(min_date = min(datetime, na.rm = TRUE),
            max_date = max(datetime, na.rm = TRUE))

# different months
tweets %>%
  mutate(month = lubridate::month(datetime)) %>%
  distinct(month)


# Explore location variable
# ------------------------------------------------------------------------------
tweets %>%
  group_by(location) %>%
  count(sort = TRUE) %>%
  ungroup %>%
  mutate(location = fct_reorder(location, n, sum),
         location = fct_lump(location, 10, w = n)) %>%
  ggplot() +
  geom_col(aes(location, n, fill = location)) +
  coord_flip()


# Top tweet's words
# ------------------------------------------------------------------------------
# Common words from tweets...
tweets %>%
  select(content) %>%
  unnest_tokens(word, content) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(word) %>%
  count(sort = TRUE) %>%
  ungroup() %>%
  slice(1:30) %>%
  mutate(word = fct_reorder(word, n, sum)) %>%
  ggplot +
  geom_col(aes(word, n, fill = word)) +
  coord_flip() +
  theme(legend.position = "none")


# Same process but using month + word as a group
tweets %>%
  mutate(month = lubridate::month(datetime)) %>%
  select(month, content) %>%
  unnest_tokens(word, content) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(month, word) %>%
  count(sort = TRUE) %>%
  ungroup() %>%
  slice(1:50) %>%
  mutate(word = fct_reorder(word, n, sum)) %>%
  ggplot +
  geom_col(aes(word, n, fill = as.factor(month))) +
  coord_flip() +
  labs(fill = "Tweet's month",
       title = "Top tweet's words by month",
       subtitle = "During Juneteenth 2021 celebration")



# using bars to represent the number of tweets per day
tweets_per_day %>%
  ggplot(aes(date, n, width = .5)) +
  geom_col(fill = "steelblue") +
  coord_polar()

# use log10 scale
tweets_per_day %>%
  ggplot(aes(date, n, width = .5)) +
  geom_col(fill = "steelblue") +
  scale_y_log10() +
  coord_polar()
```
  
```{r} 
# 2. Other geom as alternatives
# ------------------------------------------------------------------------------
# polar-tweet activty using geom_area                   
tweets_per_day %>% 
  filter(date < "2021-05-01") %>% 
  ggplot(aes(date)) +
  geom_area(aes(y=n), fill="#c51f5d",
            colour = "white",
            alpha = .50,
            size = .2) +
  scale_y_log10() +
  coord_polar() +
  labs(x = NULL,
       title="Tweet activity during Juneteenth 2021") +
  theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(size = 0.1, 
                                        arrow = arrow(length = unit(.1, "inches"))),
        plot.margin = margin(25, 15, 15, 25),
        plot.background = element_rect(color = "#FFFCFC", fill = "#141d26"),
        plot.title = element_text(colour = "white", size = 16),
        
        axis.text = element_text(colour = "white", size=12),
        axis.text.y = element_blank(),
        
        legend.title = element_text(colour = "white"),
        legend.text = element_text(colour = "white")
        )


# polar-tweet activty using geom_col
tweets_per_day %>% 
  filter(date < "2021-05-01") %>% 
   ggplot(aes(date, n, width = .5)) +
  geom_col(aes(fill = n),
           alpha = .8,
           colour = "#243447",
           size = .4) +
  scale_fill_gradient2(high = "#c51f5d") +
  scale_y_log10() +
  coord_polar() +
  labs(x = NULL,
       title="Tweet activity during Juneteenth 2021") +
  theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(size = 0.1, 
                                        arrow = arrow(length = unit(.1, "inches"))),
        plot.margin = margin(25, 15, 15, 25),
        plot.background = element_rect(color = "#FFFCFC", fill = "#141d26"),
        plot.title = element_text(colour = "white", size = 16),
        axis.text = element_text(colour = "white", size=12),
        legend.title = element_text(colour = "white"),
        legend.text = element_text(colour = "white")
        )
```






